Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-07 paper/tv.tex
m

void process (int i)
/* requires d1<d2,  emp 
   ensures  true, (Update(i)#d1.emp#d2.([x=i]Critical(i).Exit(i) \/ emp))^* */
{
  [x=-1] 
  deadline (event[Update(i)]{x := i}, d1);
  delay (d2);
  if (x=i) {
    event[Critical(i)]{ct := ct +1};
    event[Exit(i)]{ct := ct-1; x := -1};
    process (i);
  } else {
    process (i);
}

void main () 
/* requires true /\ emp 
   ensures  true, (!Critical \/ (Critical(i) /\ i=x))^* */
// ensures  true, (ct <= 1)^*
{ process(0) || process(1) || process(2); }
\end{lstlisting}  
  \vspace{0mm}
  \caption{Fischer's mutual exclusion algorithm.}\label{fig:overview_ficher}
     \vspace{0mm}
\end{figure}

@@

\begin{figure}[ht]
      \vspace{0mm}
\begin{lstlisting}[columns=fullflexible]
void addSugar(int n)
/* requires n>0,  _^*.Cup 
   ensures  t>=n, Sugar # t */
{ if n == 0 then event[Sugar] 
  else {
    timeout (oneSugar(), 1);
    addSugar (n-1);}}

void makeCoffee (int n)
/* requires n>0, emp 
 ensures  t<5/\t>=n/\t1<3, 
 Cup.(Sugar#t).(Coffee#t1).Done */
{ event[Cup];
  deadline (addSugar(n), 5);
  deadline (event[Coffee], 3);
  event[Done];}

void main () 
/* requires true, emp 
 ensures  t<8, (_^*.Done)#t */
{ makeCoffee(3);}
// make coffee with 3 lumps of sugar

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{| 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\

@@

\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Wait                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  t {\leq} d 
}{(V, \textbf{delay}[d]) \xrightarrow[]{\text{t}} (V, \textbf{delay} [d \text{-}t])}\ [de_1]} 
\qquad \qquad 
\code{\frac{
}{(V, \textbf{delay} [0]) \xrightarrow[]{\tau} (V, ())}\ [de_2]} 
\end{flalign*}

The above captures behaviors of process \code{\textbf{delay}[d]}. 
Rule \code{de_1} states that the process may idle for 
any amount of time as long as it is less than or equal to \codem{d} time units; 
Rule \code{de_2} states that the process terminates immediately after 
\codem{d} becomes \codem{0}.
\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   timeout                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e_1) \xrightarrow[]{\anyevent{A}} (V^\prime,e_1^\prime)
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\anyevent{A}} (V^\prime,e^\prime)}\ [to_1]} 
\quad  
\code{\frac{
  (V,e_1) \xrightarrow[]{\tau} (V^\prime,e_1^\prime)  
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (V^\prime, e_1^\prime\ \textbf{timeout}[d]\ e_2 )}\ [to_2]} 
\\
\code{\frac{
  (V,e_1) \xrightarrow[]{\text{t}} (V,e_1^\prime)  \quad (t {\leq} d) 
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (V,e_1^\prime\ \textbf{timeout} \  [d\text{-}t] \ e_2)}\ [to_3]} 
\quad  
\code{\frac{
}{(V, e_1\ \textbf{timeout}   [0] \ e_2) \xrightarrow[]{\tau} (V, e_2)}\ [to_4]} 
\end{flalign*}
If an observable event \anyevent{A} can be engaged by \code{e_1}, 
then \code{e_1\ \textbf{timeout}[d]\ e_2} becomes \code{e_1^\prime} (rule \code{[to_1]}). 
An invisible transition does 
not solve the choice (rule \code{[to_2]}). 
If \code{e_1} may idle for less than or equal 
to \codem{d} time units, so is the composition (rule \code{[to_3]}). 
When \codem{d} becomes \codem{0}, 
\code{e_2} takes over control by a silent transition (rule \code{[to_4]}).
\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   deadline                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e) \xrightarrow[]{\anyevent{A}} (V^\prime,e^\prime)
}{(V, \textbf{deadline} [d] \ e ) \xrightarrow[]{\anyevent{A}} (V^\prime, \textbf{deadline}  [d] \ e^\prime)}\ [dl_1]} 
\  
\code{\frac{
  (V,e) \xrightarrow[]{\text{t}} (V,e^\prime)  \quad (t {\leq} d)
}{(V, \textbf{deadline}  [d]\ e ) \xrightarrow[]{\text{t}} (V, \textbf{deadline}  [d \text{-}t] \ e^\prime )}\ [dl_2]} 
\end{flalign*}

@@

\begin{flalign*}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   deadline                 %%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \code{\frac{
    V \not\models  b 
  }{(V, [b] \ e ) \xrightarrow[]{t} (V, [b] \ e)}\ [gu_1]} 
  \quad
  \code{\frac{
    V \models  b 
  }{(V, [b] \ e ) \xrightarrow[]{\tau} (V, e)}\ [gu_2]} 
  \quad
  \code{\frac{
  }{(V, \alpha ) \xrightarrow[]{\tau} (\alpha(V), ())}\ [assign]} 
\end{flalign*}

\begin{flalign*}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   deadline                 %%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \code{\frac{
    (V,e_1) \xrightarrow[]{\anyevent{A}} (V^\prime,e_1^\prime) 
  }{(V, e_1 || e_2 ) \xrightarrow[]{\anyevent{A}} (V^\prime, e_1^\prime || e_2)}\ [pa_1]} 
  \quad  
  \code{\frac{
    (V,e_2) \xrightarrow[]{\anyevent{A}} (V^\prime,e_2^\prime) 
  }{(V, e_1 || e_2 ) \xrightarrow[]{\anyevent{A}} (V^\prime, e_1 || e_2^\prime)}\ [pa_2]} 
\quad
\code{\frac{
  (V,e_1) \xrightarrow[]{\anyevent{A}/\tau} (V^\prime,e_1^\prime) 
}{(V, e_1; e_2 ) \xrightarrow[]{\anyevent{A}/\tau} (V^\prime, e_1^\prime ; e_2)}\ [se]} 
\end{flalign*}

@@

We plant the effects specifications into the Hoare-style verification system, using \code{ \effect_{pre}} and \code{  \effect_{post} }  to capture the temporal pre/post 
condition. %\code{\effect_{pre}} and the postcondition \code{\effect_{post}}.

@@

 \m{(Timed\ Effects)}  & \ \code{\effect}   &\code{ ::=}&

   \code{ \bott
   \  | \ \empt 
    \ | \ ev 
    %   \ | \ \anyevent{a}
   \  | \ \effect_1 {\seq} \effect_2
    \ | \ \effect_1 {\vee} \effect_2
    \ | \ \effect_1 {||} \effect_2
      %    \  | \ \effect_1 \times \effect_2
      \ | \ \effect \# t 
   \  | \  \effect^\star
}

@@

The  syntax of the \timedEffects is formally defined in 
\figref{fig:Syntax_of_Types_and_Effects} Effects is 
 a conditioned time  event sequence \code{\pure{\pi},  \effect} or a disjunction of two effects $\effect_1 \vee \effect_2$.
Timed sequences comprise \textit{nil} ($\bott$);
 an empty trace $\empt$;
a single event \code{ev};
  %where \anyevent{a} is an element from a finite set of events, donated by \code{\Sigma};
%parametrized event \code{ \anyevent{a}.\iota} where \code{\iota} is a list pattern expressing a possibly infinite list (for example, \code{\anyevent{a}.[1,2,3,4]},  \code{\anyevent{a}.[1 .. 4]} and \code{\anyevent{a}.[1, 2.. 4]} all refer to the finite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot \anyevent{a}.4},
%while \code{\anyevent{a}.[1 ..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot}..., and \code{\anyevent{a}.[1, 3..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.3 \cdot \anyevent{a}.5\cdot}...);
 sequences concatenation \code{\effect_1\cdot \effect_2};
disjunction  \code{\effect_1\vee \effect_2};
parallel composition of two processes is written as \code{\effect_1 || \effect_2}, where \code{\effect_1} and \code{\effect_2} 
may communicate via multi-party event synchronization or shared variables. 
We introduce a new operator \code{\#}, and the effects \code{\effect \# t } represents that a trace takes real-time \code{t} to complete, where \code{t} is a \emph{term}. 
A time sequence can be constructed by \code{\star}, representing zero or more times repetition of a trace.

@@

To define the model, 
\code{var} is the set of program variables, 
\code{val} is the set of primitive values, 
\code{\effect} is the set of event sequences (or event multi-trees, per se), indicating the sequencing constraints on temporal behaviour.
Let \code{d, s, \varphi \models \effect} denote the model relation, i.e., 
the effects \code{\effect} take exactly \code{d} time units to complete; 
the stack \code{s} and linear temporal events \code{\varphi} satisfy the 
temporal effects \code{\effect}, with \code{d}, \code{s}, \code{\varphi} 
from the following concrete domains: \code{d}  {\defeq}\  \code{\mathbb{N}}, 
\code{s}  \defeq\  \code{var {\rightarrow} val } and \code{\varphi}   \defeq\ \code{ev^*} \ .

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.2}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models \pure{\pi}  , \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} }\ and \ \varphi {=} [ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, (\effect_1\cdot\effect_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \\
&&& \qquad  s.t. \ \code{d_1, s, \varphi_1 {\models} \pure{\pi}  , \effect_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} \pure{\pi}  , \effect_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , (\effect_1{\vee}\effect_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , \effect_1 } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , \effect_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev_1{\cdot}\effect_1)||(ev_2{\cdot}\effect_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , ev_1{\cdot} (\effect_1||(ev_2{\cdot}\effect_2)) } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , ev_2{\cdot}((ev_1{\cdot}\effect_1)||\effect_2)  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel  2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev\#t_1) || (\epsilon \# t_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {\geq}t_2)}, (ev\#t_1) } 
\ or\ \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {<}t_2)}  , (ev\#t_1){\cdot}(\epsilon \# (t_2\text{-}t_1))  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  ,\effect\# t}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  , t {=} 0 \rrbracket_{s} {=}  
 \pure{True} }, \ \code{d, s, \varphi \models \pure{\pi}  , \empt}\ or  \\
& & &  \code{\llbracket \pure{\pi}, t {>} 0 \rrbracket_{s} {=}  
\pure{True} }, 
\  \exists \code{\effect_1,\effect_2}. \  \code{\effect_1 {\cdot} \effect_2 {=} \effect}\\
&&& \qquad  s.t. \ \code{d, s, \varphi {\models} \pure{(\pi 
{\wedge} t_1{\geq}0{\wedge}t_2{\geq}0 {\wedge} t_1{+} t_2{=}t)}  , 
(\effect_1\#t_1) {\cdot} (\effect_2\#t_2) } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    star         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, \effect^\star}  &\m{iff}\quad & 
\code{d, s, \varphi \models \pure{\pi}  , \empt}\ or \ 
\code{d, s, \varphi \models \pure{\pi}  , \effect {\cdot} \effect^\star}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    FALSE         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  \rrbracket_{s} {=} \pure{False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}

@@

\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \effect,\Pi \}\  v \ 
\{ \effect, \Pi \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  }
{  \vdash \{  \effect, \Pi \}\  [b]\ e \ 
\{ \effect, \Pi{\wedge} b  \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}

>>>>

void process (int i)
/* req  d1<d2,  emp 
   ens  true, (Update(i)#d1.emp#d2.([x=i]Critical(i).Exit(i) \/ emp))^* */
{
  [x=-1] 
  deadline (event[Update(i)]{x := i}, d1);
  delay (d2);
  if (x=i) {
    event[Critical(i)]{ct := ct +1};
    event[Exit(i)]{ct := ct-1; x := -1};
    process (i);
  } else {
    process (i);
}

void main () 
/* req true /\ emp 
   ens  true, (!Critical \/ (Critical(i) /\ i=x))^* */
// ens  true, (ct <= 1)^*
{ process(0) || process(1) || process(2); }
\end{lstlisting}  
  \vspace{0mm}
  \caption{Fischer's mutual exclusion algorithm.}\label{fig:overview_ficher}
     \vspace{0mm}
\end{figure}

@@

\begin{figure}[ht]
      \vspace{0mm}
\begin{lstlisting}[columns=fullflexible]
void addSugar(int n)
/* req n>0,  _^*.Cup 
   ens  t>=n, Sugar # t */
{ if n == 0 then event[Sugar] 
  else {
    timeout (oneSugar(), 1);
    addSugar (n-1);}}

void makeCoffee (int n)
/* req n>0, emp 
 ens  t<5/\t>=n/\t1<3, 
 Cup.(Sugar#t).(Coffee#t1).Done */
{ event[Cup];
  deadline (addSugar(n), 5);
  deadline (event[Coffee], 3);
  event[Done];}

void main () 
/* req true, emp 
 ens  t<8, (_^*.Done)#t */
{ makeCoffee(3);}
// make coffee with 3 lumps of sugar

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{| 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\

@@

{{\small
\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Wait                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  t {\leq} d 
}{(V, \textbf{delay}[d]) \xrightarrow[]{\text{t}} (V, \textbf{delay} [d \text{-}t])}\ [de_1]} 
\qquad \qquad 
\code{\frac{
}{(V, \textbf{delay} [0]) \xrightarrow[]{\tau} (V, ())}\ [de_2]} 
\end{flalign*}}}

The above captures behaviors of process \code{\textbf{delay}[d]}. 
Rule \code{de_1} states that the process may idle for 
any amount of time as long as it is less than or equal to \codem{d} time units; 
Rule \code{de_2} states that the process terminates immediately after 
\codem{d} becomes \codem{0}.
{{\small\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   timeout                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e_1) \xrightarrow[]{\anyevent{A}} (V^\prime,e_1^\prime)
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\anyevent{A}} (V^\prime,e^\prime)}\ [to_1]} 
\quad  
\code{\frac{
  (V,e_1) \xrightarrow[]{\tau} (V^\prime,e_1^\prime)  
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (V^\prime, e_1^\prime\ \textbf{timeout}[d]\ e_2 )}\ [to_2]} 
\\
\code{\frac{
  (V,e_1) \xrightarrow[]{\text{t}} (V,e_1^\prime)  \quad (t {\leq} d) 
}{(V, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (V,e_1^\prime\ \textbf{timeout} \  [d\text{-}t] \ e_2)}\ [to_3]} 
\quad  
\code{\frac{
}{(V, e_1\ \textbf{timeout}   [0] \ e_2) \xrightarrow[]{\tau} (V, e_2)}\ [to_4]} 
\end{flalign*}}}
If an observable event \anyevent{A} can be engaged by \code{e_1}, 
then \code{e_1\ \textbf{timeout}[d]\ e_2} becomes \code{e_1^\prime} (rule \code{[to_1]}). 
An invisible transition does 
not solve the choice (rule \code{[to_2]}). 
If \code{e_1} may idle for less than or equal 
to \codem{d} time units, so is the composition (rule \code{[to_3]}). 
When \codem{d} becomes \codem{0}, 
\code{e_2} takes over control by a silent transition (rule \code{[to_4]}).
{{\small\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   deadline                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e) \xrightarrow[]{\anyevent{A}} (V^\prime,e^\prime)
}{(V, \textbf{deadline} [d] \ e ) \xrightarrow[]{\anyevent{A}} (V^\prime, \textbf{deadline}  [d] \ e^\prime)} [dl_1]} 
\quad 
\code{\frac{
  (V,e) \xrightarrow[]{\text{t}} (V,e^\prime)  \quad (t {\leq} d)
}{(V, \textbf{deadline}  [d]\ e ) \xrightarrow[]{\text{t}} (V, \textbf{deadline}  [d \text{-}t] \ e^\prime )} [dl_2]} 
\end{flalign*}}}

@@

{{\small\begin{flalign*}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   deadline                 %%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \code{\frac{
    V \not\models  b 
  }{(V, [b] \ e ) \xrightarrow[]{t} (V, [b] \ e)}\ [gu_1]} 
  \ \ \qquad
  \code{\frac{
    V \models  b 
  }{(V, [b] \ e ) \xrightarrow[]{\tau} (V, e)}\ [gu_2]} 
  \ \ \qquad 
  \code{\frac{
  }{(V, \alpha ) \xrightarrow[]{\tau} (\alpha(V), ())}\ [assign]} 
\end{flalign*}}}

{{\small\begin{flalign*}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   deadline                 %%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \code{\frac{
    (V,e_1) \xrightarrow[]{\anyevent{A}} (V^\prime,e_1^\prime) 
  }{(V, e_1 || e_2 ) \xrightarrow[]{\anyevent{A}} (V^\prime, e_1^\prime || e_2)}\ [pa_1]} 
  \ \   
  \code{\frac{
    (V,e_2) \xrightarrow[]{\anyevent{A}} (V^\prime,e_2^\prime) 
  }{(V, e_1 || e_2 ) \xrightarrow[]{\anyevent{A}} (V^\prime, e_1 || e_2^\prime)}\ [pa_2]} 
\ \ 
\code{\frac{
  (V,e_1) \xrightarrow[]{\anyevent{A}/\tau} (V^\prime,e_1^\prime) 
}{(V, e_1; e_2 ) \xrightarrow[]{\anyevent{A}/\tau} (V^\prime, e_1^\prime ; e_2)}\ [se]} 
\end{flalign*}}}

@@

We plant the effects specifications into the Hoare-style verification system, using \code{ es_{pre}} and \code{  es_{post} }  to capture the temporal pre/post 
condition. %\code{es_{pre}} and the postcondition \code{es_{post}}.

@@

    \m{(Timed\ Effects)} &  \effect &\code{  ::= }&
    \code{
          (\pure{\pi}, es)
   \ | \ \effect_1 \vee \effect_2
}
    \\
 \m{(Event \ Sequences)}  & \ \code{es}   &\code{ ::=}&

   \code{ \bott
   \  | \ \empt 
    \ | \ ev 
    %   \ | \ \anyevent{a}
   \  | \ es_1 {\seq} es_2
    \ | \ es_1 {\vee} es_2
    \ | \ es_1 {||} es_2
      %    \  | \ es_1 \times es_2
      \ | \ es \# t 
   \  | \  es^\star
}

@@

The  syntax of the \timedEffects is formally defined in 
\figref{fig:Syntax_of_Types_and_Effects} Effects is 
 a conditioned time  event sequence \code{\pure{\pi},  es} or a disjunction of two effects $es_1 \vee es_2$.
Timed sequences comprise \textit{nil} ($\bott$);
 an empty trace $\empt$;
a single event \code{ev};
  %where \anyevent{a} is an element from a finite set of events, donated by \code{\Sigma};
%parametrized event \code{ \anyevent{a}.\iota} where \code{\iota} is a list pattern expressing a possibly infinite list (for example, \code{\anyevent{a}.[1,2,3,4]},  \code{\anyevent{a}.[1 .. 4]} and \code{\anyevent{a}.[1, 2.. 4]} all refer to the finite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot \anyevent{a}.4},
%while \code{\anyevent{a}.[1 ..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot}..., and \code{\anyevent{a}.[1, 3..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.3 \cdot \anyevent{a}.5\cdot}...);
 sequences concatenation \code{es_1\cdot es_2};
disjunction  \code{es_1\vee es_2};
parallel composition of two processes is written as \code{es_1 || es_2}, where \code{es_1} and \code{es_2} 
may communicate via multi-party event synchronization or shared variables. 
We introduce a new operator \code{\#}, and the effects \code{es \# t } represents that a trace takes real-time \code{t} to complete, where \code{t} is a \emph{term}. 
A time sequence can be constructed by \code{\star}, representing zero or more times repetition of a trace.

@@

To define the model, 
\code{var} is the set of program variables, 
\code{val} is the set of primitive values, 
\code{es} is the set of event sequences (or event multi-trees, per se), indicating the sequencing constraints on temporal behaviour.
Let \code{d, s, \varphi \models es} denote the model relation, i.e., 
the effects \code{es} take exactly \code{d} time units to complete; 
the stack \code{s} and linear temporal events \code{\varphi} satisfy the 
temporal effects \code{es}, with \code{d}, \code{s}, \code{\varphi} 
from the following concrete domains: \code{d}  {\defeq}\  \code{\mathbb{N}}, 
\code{s}  \defeq\  \code{var {\rightarrow} val } and \code{\varphi}   \defeq\ \code{ev^*} \ .

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.2}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models \pure{\pi}  , \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} }\ and \ \varphi {=} [ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, (es_1\cdot es_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \\
&&& \qquad  s.t. \ \code{d_1, s, \varphi_1 {\models} \pure{\pi}  , es_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} \pure{\pi}  , es_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , (es_1{\vee}es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , es_1 } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , es_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev_1{\cdot}es_1)||(ev_2{\cdot}es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , ev_1{\cdot} (es_1||(ev_2{\cdot}es_2)) } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , ev_2{\cdot}((ev_1{\cdot}es_1)||es_2)  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel  2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev\#t_1) || (\epsilon \# t_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {\geq}t_2)}, (ev\#t_1) } 
\ or\ \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {<}t_2)}  , (ev\#t_1){\cdot}(\epsilon \# (t_2\text{-}t_1))  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  ,es\# t}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  , t {=} 0 \rrbracket_{s} {=}  
 \pure{True} }, \ \code{d, s, \varphi \models \pure{\pi}  , \empt}\ or  \\
& & &  \code{\llbracket \pure{\pi}, t {>} 0 \rrbracket_{s} {=}  
\pure{True} }, 
\  \exists \code{es_1,es_2}. \  \code{es_1 {\cdot} es_2 {=} es}\\
&&& \qquad  s.t. \ \code{d, s, \varphi {\models} \pure{(\pi 
{\wedge} t_1{\geq}0{\wedge}t_2{\geq}0 {\wedge} t_1{+} t_2{=}t)}  , 
(es_1\#t_1) {\cdot} (es_2\#t_2) } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    star         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, es^\star}  &\m{iff}\quad & 
\code{d, s, \varphi \models \pure{\pi}  , \empt}\ or \ 
\code{d, s, \varphi \models \pure{\pi}  , es {\cdot} es^\star}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    FALSE         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  \rrbracket_{s} {=} \pure{False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}

@@

{{\small\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \effect,\Pi \}\  v \ 
\{ \effect, \Pi \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  }
{  \vdash \{  \effect, \Pi \}\  [b]\ e \ 
\{ \effect, \Pi{\wedge} b  \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}}}
